Intelligent Reflecting Surface-Aided SWIPT: Joint Waveform, Active and Passive Beamforming Design

### 2020-11-04 main.tex
:construction: Rephrase and revise

	\begin{abstract}
		The performance of Simultaneous Wireless Information and Power Transfer (SWIPT) is mainly restricted by the strength of the received Radio-Frequency (RF) signal. To tackle this problem, we introduce a low-power Intelligent Reflecting Surface (IRS) that compensates the propagation loss and boosts the energy efficiency by provide a passive beamforming gain. This paper investigates an IRS-aided SWIPT system where a multi-carrier multi-antenna Access Point (AP) transmits information and power simultaneously to a single-antenna user under the assist of an IRS. Considering the rectifier nonlinearity, we aim to maximize the Rate-Energy (R-E) region through a joint optimization of the transmit waveform and active beamforming at the AP, the reflection coefficients at the IRS, and the power splitting ratio at the user. Stationary solutions are achieved based on Alternating Optimization (AO) technique, where the optimal active beamforming is obtained in closed form, the passive beamforming is optimized by Successive Convex Approximation (SCA) technique, and the waveform and splitting ratio are optimized by Geometric Programming (GP) technique. It is confirmed that due to rectifier nonlinearity, a dedicated power signal can be beneficial to energy harvesting (EH), such that the Time-Switching (TS) receiver is preferred over Power-Splitting (PS) receiver for multi-carrier transmission at a low Signal-to-Noise Ratio (SNR). Simulation results also show that the introduction of IRS significantly enlarges the R-E region while preserving the waveform preference and transceiving strategy. Although the AP-IRS-user extra link is relatively weak, we found that the information and power benefits of passive beamforming scale much faster than those of active beamforming. It suggests using more active antennas for small-scale SWIPT and more passive elements for large-scale SWIPT.
	\end{abstract}

@@

	\begin{section}{Problem Formulation}\label{se:problem_formulation}
		We characterize the R-E region through a current maximization problem subject to transmit power, IRS magnitude, and rate constraints
		\begin{maxi!}
			{\boldsymbol{\phi},\boldsymbol{w}_I,\boldsymbol{w}_P,\rho}{z(\boldsymbol{\phi},\boldsymbol{w}_I,\boldsymbol{w}_P,\rho)}{\label{op:original}}{\label{ob:original}}
			\addConstraint{\frac{1}{2}\left(\lVert{\boldsymbol{w}_I}\rVert^2+\lVert{\boldsymbol{w}_P}\rVert^2\right)\le{P}}\label{co:original_power}
			\addConstraint{R(\boldsymbol{\phi},\boldsymbol{w}_I,\rho) \ge \bar{R}}\label{co:original_rate}
			\addConstraint{\lvert{\phi_l}\rvert=1, \quad l=1,\dots,L}\label{co:original_modulus}
			\addConstraint{0 \le \rho \le 1\,.}
		\end{maxi!}
		Problem~\eqref{op:original} is intricate as the variables are coupled and the objective function \eqref{ob:original}, the rate constraint \eqref{co:original_rate}, and the IRS magnitude constraint \eqref{co:original_modulus} are non-convex. To reduce the design complexity, we propose a suboptimal AO algorithm that iteratively updates the waveform and active beamforming at the transmitter, the phase shift at the IRS, and the power splitting ratio at the receiver, until convergence.

@@

			where $\boldsymbol{V}=[\boldsymbol{V}_1,\dots,\boldsymbol{V}_N] \in \mathbb{C}^{L \times MN}$, $\boldsymbol{M}=[\boldsymbol{V}^H, \boldsymbol{h}_{D}]^H \in \mathbb{C}^{(L+1) \times MN}$, $\boldsymbol{C}_{I/P,k}=\boldsymbol{M}\boldsymbol{W}_{I/P,k}\boldsymbol{M}^H \in \mathbb{C}^{(L+1)\times(L+1)}$. Therefore, the rate and objective expressions rewrite as

@@

			Problem~\eqref{op:irs} is not a standard Semidefinite Programming (SDP). If we relax the rank constraint \eqref{co:irs_rank} to formulate a convex problem, there is no guarantee $\boldsymbol{\Phi}^{\star}$ is rank-\num{1}, and the optimal rank-\num{1} solution $\bar{\boldsymbol{\phi}}^{\star}$ extracted from $\boldsymbol{\Phi}^{\star}$ may not be a stationary point of the original problem~\eqref{op:original}. In Section~\ref{se:performance_evaluation}, we numerically show that $\boldsymbol{\Phi}^{\star}$ is rank-\num{1} for all tested channel realizations under different configurations such that the performance loss by SDR is insignificant. A related version of problem~\eqref{op:irs} can be solved using existing optimization tools such as CVX \cite{Grant2013}.

@@

			The denominators of \eqref{co:waveform_objective}, \eqref{co:waveform_rate} are posynomials \cite{Boyd2007} that further decompose as

@@

		\begin{subsection}{Convergence}
			\begin{proposition}\label{pr:irs}
				For any feasible initial point, the proposed SCA-based Algorithm~\ref{al:irs} can provide a feasible $\boldsymbol{\Phi}^{\star}$ that satisfy the KKT conditions, although there is no guarantee $\boldsymbol{\Phi}^{\star}$ is rank-\num{1}.
			\end{proposition}

			\begin{proof}\label{pf:irs}
				The objective function \eqref{ob:irs} is non-decreasing over iterations because the solution of problem~\eqref{op:irs} at iteration $i-1$ is still a feasible point at iteration $i$. Moreover, the sequence $\{\tilde{z}(\boldsymbol{\Phi}^{(i)})\}_{i=1}^{\infty}$ is bounded above due to the unit-modulus constraint \eqref{co:irs_modulus}. Thus, Algorithm~\ref{al:irs} is guaranteed to converge. To prove $\boldsymbol{\Phi}^{(i)}$ converge to the set of stationary points of IRS subproblem, we notice that the SCA-based Algorithm~\ref{al:irs} is indeed an inner approximation algorithm \cite{Marks1978}, since $\tilde{z}(\boldsymbol{\Phi}) \le z(\boldsymbol{\Phi})$, $\partial\tilde{z}(\boldsymbol{\Phi}^{(i)})/\partial\boldsymbol{\Phi}=\partial z(\boldsymbol{\Phi}^{(i)})/\partial\boldsymbol{\Phi}$ and the approximation \eqref{eq:taylor_1} -- \eqref{eq:taylor_3} are asymptotically tight as $i \to \infty$ \cite{Li2013}. Therefore, Algorithm~\ref{al:irs} is guaranteed to provide a feasible $\boldsymbol{\Phi}^{\star}$ that satisfy the KKT conditions.
			\end{proof}

@@

			\begin{proposition}\label{pr:ao}
				Every limit point $(\boldsymbol{\phi}^{\star},\boldsymbol{w}_I^{\star},\boldsymbol{w}_P^{\star},\rho^{\star})$ of the proposed alternating algorithm is a stationary point of the original problem~\eqref{op:original}.
			\end{proposition}

@@

	\begin{section}{Performance Evaluations}\label{se:performance_evaluation}
		\begin{figure}[!t]
			\centering
			\def\svgwidth{\columnwidth}
			\import{assets/}{layout.pdf_tex}
			\caption{System layout in the simulation.}
			\label{fi:layout}
		\end{figure}
		To evaluate the performance of the proposed IRS-aided SWIPT system, we characterize the average R-E regions under typical setups. Consider a large open space WiFi-like environment at a center frequency of \SI{5.18}{\GHz} with reference bandwidth $B=1\,\si{\MHz}$. As shown in Fig.~\ref{fi:layout}, we assume the IRS moves along a horizontal line parallel to the AP-user path and let $d_H$, $d_V$ be the horizontal and vertical distances from the AP to the IRS, respectively. We also denote $d_D$, $d_I$, $d_R$ as the length of direct, incident, reflective paths such that $d_I=\sqrt{d_H^2+d_V^2}$, $d_R=\sqrt{(d_D-d_H)^2+d_V^2}$. Choose as reference $d_D=15\,\si{\meter}$, $d_V=2\,\si{\meter}$ and $d_H=2\,\si{\meter}$. The path loss and fading parameters are obtained from IEEE TGn channel model D \cite{Erceg2004}, and reference path loss is set to $L_0=-35\,\si{\dB}$ at $d_0=1\,\si{\meter}$. We assume all channels are NLoS with taps modelled as i.i.d. CSCG random variables of unit average sum-power to create a normalized multipath response. The reference numbers of transmit antennas, IRS reflectors and subbands are $M=1$, $L=20$, $N=16$, respectively. No spatial correlation is considered across transmit antennas and IRS arrays. Rectenna parameters are taken as $k_2=0.0034$, $k_4=0.3829$, $R_{\text{ant}}=50\,\si{\ohm}$. With \SI{0}{\dBi} transmit antenna gain, the average Effective Isotropic Radiated Power (EIRP) is fixed to $P=-36\,\si{\dBm}$ while the reference average noise power is $\sigma_n=-40\,\si{\dBm}$ at all subbands. We also assume \SI{0}{\dBi} IRS element gain and \SI{2}{\dBi} receive antenna gain. For the algorithm, the tolerance is $\epsilon=10^{-8}$, the number of candidates in the Gaussian randomization method is $Q=10^{3}$, and the R-E region is averaged over \num{200} channel realizations. In the R-E boundary, the leftmost point corresponds to WPT ($\rho=1$) where power can be allocated simultaneously to modulated and unmodulated waveform to maximize the average output DC current. On the other hand, the rightmost point corresponds to WIT ($\rho=0$) where the solution coincides with the Water-Filling (WF) algorithm that allocates all power to modulated waveform only. For a fair comparison, the $x$-axis of the plots has been normalized to per-subband rate.

>>>>

	\begin{abstract}
		The performance of Simultaneous Wireless Information and Power Transfer (SWIPT) is mainly restricted by the strength of the received Radio-Frequency (RF) signal. To tackle this problem, we introduce a low-power Intelligent Reflecting Surface (IRS) that compensates the propagation loss and boosts the energy efficiency by provide a passive beamforming gain. This paper investigates an IRS-aided SWIPT system where a multi-carrier multi-antenna Access Point (AP) transmits information and power simultaneously to a single-antenna user under the assist of an IRS. Considering the rectifier nonlinearity, we aim to maximize the Rate-Energy (R-E) region through a joint optimization of the transmit waveform and active beamforming at the AP, the reflection coefficients at the IRS, and the power splitting ratio at the user. Stationary solutions are achieved based on Alternating Optimization (AO) technique, where the optimal active beamforming is obtained in closed form, the passive beamforming is optimized by Successive Convex Approximation (SCA) technique, and the waveform and splitting ratio are optimized by Geometric Programming (GP) technique. It is confirmed that due to rectifier nonlinearity, a dedicated power signal can be beneficial to energy harvesting (EH), such that the Time-Switching (TS) receiver is preferred over Power-Splitting (PS) receiver for multi-carrier transmission at a low Signal-to-Noise Ratio (SNR). Simulation results also show that the introduction of IRS significantly enlarges the R-E region while preserving the waveform preference and transceiving strategy. Although the AP-IRS-user extra link is relatively weak, we found that the information and power benefits of passive beamforming scale much faster than those of active beamforming. It suggests using more active antennas for small-scale SWIPT and more passive elements for large-scale SWIPT. % The proposed design enables a joint control of transmitter, receiver and wireless channel.
	\end{abstract}

@@

	\begin{section}{Problem Formulation}\label{se:problem_formulation}
		We characterize the R-E region through a current maximization problem subject to transmit power, IRS magnitude, and rate constraints
		\begin{maxi!}
			{\boldsymbol{\phi},\boldsymbol{w}_I,\boldsymbol{w}_P,\rho}{z(\boldsymbol{\phi},\boldsymbol{w}_I,\boldsymbol{w}_P,\rho)}{\label{op:original}}{\label{ob:original}}
			\addConstraint{\frac{1}{2}\left(\lVert{\boldsymbol{w}_I}\rVert^2+\lVert{\boldsymbol{w}_P}\rVert^2\right)\le{P}}\label{co:original_power}
			\addConstraint{R(\boldsymbol{\phi},\boldsymbol{w}_I,\rho) \ge \bar{R}}\label{co:original_rate}
			\addConstraint{\lvert{\phi_l}\rvert=1, \quad l=1,\dots,L}\label{co:original_modulus}
			\addConstraint{0 \le \rho \le 1\,.}
		\end{maxi!}
		Problem~\eqref{op:original} is intricate as the variables are coupled and the objective function \eqref{ob:original}, the rate constraint \eqref{co:original_rate}, and the IRS magnitude constraint \eqref{co:original_modulus} are non-convex. To achieve a feasible solution, we propose a AO algorithm that iteratively updates the waveform and active beamforming at the transmitter, the phase shift at the IRS, and the power splitting ratio at the receiver, until convergence.
		\begin{remark}
			At subband $n$, the MIMO incident channel $\boldsymbol{H}_{I,n}$ can be divided into $L$ MISO channels between all transmit antennas and each IRS element. Reflection coefficient $\phi_l$ is shared for the $l$-th MISO incoming channel of all subbands, namely $MN$ entries in total. As information and power transfer favor different channel strength distribution over frequencies, for a given channel, the passive beamforming need to be optimized for each R-E point to maximize the R-E tradeoff. If the IRS is frequency-selective and there is only one transmit antenna, then reflector $l$ can simply align the AP-IRS($l$)-user channel with the AP-user channel at all subbands so that the joint design of waveform, active and passive beamforming, and splitting ratio would follow straightforwardly from existing SWIPT work \cite{Clerckx2018b}.
		\end{remark}

@@

			where $\boldsymbol{V}=[\boldsymbol{V}_1,\dots,\boldsymbol{V}_N] \in \mathbb{C}^{L \times MN}$, $\boldsymbol{M}=[\boldsymbol{V}^H, \boldsymbol{h}_{D}]^H \in \mathbb{C}^{(L+1) \times MN}$, $\boldsymbol{C}_{I/P,k}=\boldsymbol{M}\boldsymbol{W}_{I/P,k}\boldsymbol{M}^H \in \mathbb{C}^{(L+1)\times(L+1)}$. Therefore, we rewrite the rate and objective expressions as

@@

			Problem~\eqref{op:irs} is not a standard Semidefinite Programming (SDP). If we relax the rank constraint \eqref{co:irs_rank} to formulate a convex problem, there is no guarantee $\boldsymbol{\Phi}^{\star}$ is rank-\num{1}, and the optimal rank-\num{1} solution $\bar{\boldsymbol{\phi}}^{\star}$ extracted from $\boldsymbol{\Phi}^{\star}$ may not be a stationary point of the original problem~\eqref{op:original}. In Section~\ref{se:performance_evaluation}, we numerically show that $\boldsymbol{\Phi}^{\star}$ is rank-\num{1} for all tested channel realizations under different configurations such that the performance loss by SDR is insignificant. Problem~\eqref{op:irs} without constraint \ref{co:irs_rank} can be solved using existing optimization tools such as CVX \cite{Grant2013}.

@@

			The denominators of \eqref{co:waveform_objective}, \eqref{co:waveform_rate} are posynomials \cite{Boyd2007} that are further decomposed as

@@

		\begin{subsection}{Convergence}
			\begin{proposition}\label{pr:irs}
				For any feasible initial point, the proposed SCA-based Algorithm~\ref{al:irs} can provide a feasible $\boldsymbol{\Phi}^{\star}$ that satisfies the KKT conditions, although there is no guarantee $\boldsymbol{\Phi}^{\star}$ is rank-\num{1}.
			\end{proposition}

			\begin{proof}\label{pf:irs}
				The objective function \eqref{ob:irs} is non-decreasing over iterations because the solution of problem~\eqref{op:irs} at iteration $i-1$ is still a feasible point at iteration $i$. Moreover, the sequence $\{\tilde{z}(\boldsymbol{\Phi}^{(i)})\}_{i=1}^{\infty}$ is bounded above due to the unit-modulus constraint \eqref{co:irs_modulus}. Thus, Algorithm~\ref{al:irs} is guaranteed to converge. To prove $\boldsymbol{\Phi}^{(i)}$ converges to the set of stationary points of IRS subproblem, we notice that the SCA-based Algorithm~\ref{al:irs} is indeed an inner approximation algorithm \cite{Marks1978}, since $\tilde{z}(\boldsymbol{\Phi}) \le z(\boldsymbol{\Phi})$, $\partial\tilde{z}(\boldsymbol{\Phi}^{(i)})/\partial\boldsymbol{\Phi}=\partial z(\boldsymbol{\Phi}^{(i)})/\partial\boldsymbol{\Phi}$ and the approximation \eqref{eq:taylor_1} -- \eqref{eq:taylor_3} are asymptotically tight as $i \to \infty$ \cite{Li2013}. Therefore, Algorithm~\ref{al:irs} is guaranteed to provide a feasible $\boldsymbol{\Phi}^{\star}$ that satisfies the KKT conditions.
			\end{proof}

@@

			\begin{proposition}\label{pr:ao}
				Suppose $\boldsymbol{\Phi}^{\star}$ is rank-\num{1}, every limit point $(\boldsymbol{\phi}^{\star},\boldsymbol{w}_I^{\star},\boldsymbol{w}_P^{\star},\rho^{\star})$ of the proposed alternating algorithm is a stationary point of the original problem~\eqref{op:original}.
			\end{proposition}

@@

	\begin{section}{Performance Evaluations}\label{se:performance_evaluation}
		\begin{figure}[!t]
			\centering
			\def\svgwidth{\columnwidth}
			\import{assets/}{layout.pdf_tex}
			\caption{System layout in the simulation.}
			\label{fi:layout}
		\end{figure}
		To evaluate the performance of the proposed IRS-aided SWIPT system, we characterize the average R-E regions under typical setups. Consider a large open space WiFi-like environment at a center frequency of \SI{5.18}{\GHz} with reference bandwidth $B=1\,\si{\MHz}$. As shown in Fig.~\ref{fi:layout}, we assume the IRS moves along a horizontal line parallel to the AP-user path and let $d_H$, $d_V$ be the horizontal and vertical distances from the AP to the IRS, respectively. We also denote $d_D$, $d_I$, $d_R$ as the length of direct, incident, reflective paths such that $d_I=\sqrt{d_H^2+d_V^2}$, $d_R=\sqrt{(d_D-d_H)^2+d_V^2}$. We choose as reference $d_D=15\,\si{\meter}$, $d_V=2\,\si{\meter}$ and $d_H=2\,\si{\meter}$. The path loss and fading parameters are obtained from IEEE TGn channel model D \cite{Erceg2004}, and reference path loss is set to $L_0=-35\,\si{\dB}$ at $d_0=1\,\si{\meter}$. We assume all channels are NLoS with taps modelled as i.i.d. CSCG random variables of unit average sum-power to create a normalized multipath response. The reference numbers of transmit antennas, IRS reflectors and subbands are $M=1$, $L=20$, $N=16$, respectively. No spatial correlation is considered across transmit antennas and IRS arrays. Rectenna parameters are taken as $k_2=0.0034$, $k_4=0.3829$, $R_{\text{ant}}=50\,\si{\ohm}$. With \SI{0}{\dBi} transmit antenna gain, the average Effective Isotropic Radiated Power (EIRP) is fixed to $P=-36\,\si{\dBm}$ while the reference average noise power is $\sigma_n=-40\,\si{\dBm}$ at all subbands. We also assume \SI{0}{\dBi} IRS element gain and \SI{2}{\dBi} receive antenna gain. For the algorithm, the tolerance is $\epsilon=10^{-8}$, the number of candidates in the Gaussian randomization method is $Q=10^{3}$, and the R-E region is averaged over \num{200} channel realizations. In the R-E boundary, the leftmost point corresponds to WPT ($\rho=1$) where power can be allocated simultaneously to modulated and unmodulated waveform to maximize the average output DC current.
		% - at the WPT point, power is allocated to the multisine waveform only, isn't it?
		% - for single-carrier WPT, modulated waveform may be preferred as it has a modulation gain (also proved in Fig. 5)
		On the other hand, the rightmost point corresponds to WIT ($\rho=0$) where the solution coincides with the Water-Filling (WF) algorithm that allocates all power to modulated waveform only. For a fair comparison, the $x$-axis of the plots has been normalized to per-subband rate.
